---
title: "Lab 02 - Plastic waste"
author: "Conor Lacey"
date: "01/19/23"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

Histogram of plastic waste per capita faceted by continent

```{r plastic-waste-continent}
plastic_waste %>%filter(plastic_waste_per_cap < 3.5) %>% #filtering out 3.5
ggplot(aes(x = plastic_waste_per_cap)) +
  facet_wrap(~continent)+
  labs(title = "Plastic Waste Per Capita by Continent")+
  xlab("Plastic Waste Per Capita")+
  ylab("Count")+
  geom_histogram(binwidth = 0.1,color="black")
```

From the histogram it appears Africa is the least variant continent
of all the continents. It also appears that North America is the most variant 
continent of all the continents.

### Exercise 2

The density distribution of plastic waste per capita faceted by continent
```{r plastic-waste-density}
plastic_waste %>%filter(plastic_waste_per_cap < 3.5) %>% 
ggplot(aes(x = plastic_waste_per_cap,
           color=continent,
           fill=continent)) +
  labs(title ="Plastic waste per capita by continent")+
  xlab("Plastic Waste Per Capita")+
  ylab("Density")+
  geom_density(alpha=0.3)
```

### Exercise 3

We defined the color and fill of the curves by mapping the aesthetics because they are
based on the data values (the continent label). The alpha level is not based on the data values,therefore it goes in the geom as a characteristic. 

### Exercise 4

In contrast to box plots, violins reveal the continuous density. It is apparent here
that the violins are just a mirrored density distribution of each continent displayed 
in a similar way to box plots. 

The box plots are able to display the quartiles and the median of the data. Violins 
cannot do this.

```{r plastic-waste-violin}
plastic_waste %>%filter(plastic_waste_per_cap < 3.5) %>% 
ggplot( mapping = aes(x = continent, 
                      y = plastic_waste_per_cap)) +
    labs(title ="Plastic waste per capita by continent")+
  ylab("Plastic Waste Per Capita")+
  xlab("Continent")+
  geom_violin()
```

### Exercise 5

There appears to be a positive relationship here between plastic and mismanaged plastic 
waste per capita. However, it also appears this relationship weakens with increased levels of mismanaged plastic waste per capita.

```{r plastic-waste-mismanaged}
plastic_waste %>%filter(plastic_waste_per_cap < 3.5) %>% 
ggplot( mapping = aes(x = mismanaged_plastic_waste_per_cap, 
                      y =plastic_waste_per_cap)) +
    labs(title ="Relationship between Plastic and Mismanaged Plastic Per Capita")+
  ylab("Plastic Waste Per Capita")+
  xlab("Mismanaged Plastic Waste Per Captia")+
  geom_point()
```

### Exercise 6

Coloring the points by continent it appears there is a positive relationship 
between plastic waste per capita and mismanaged plastic waster per capita.The
relationship is still weird however when looking at Asia as the relationship weakens
as the level of mismanaged plastic waste per capita increases. There also appears to 
be no relationship in Europe.

```{r plastic-waste-mismanaged-continent}
plastic_waste %>%filter(plastic_waste_per_cap < 3.5) %>% 
ggplot( mapping = aes(x = mismanaged_plastic_waste_per_cap, 
                      y =plastic_waste_per_cap,
                      color=continent)) +
    labs(title ="Relationship between Plastic and Mismanaged Plastic Per Capita")+
  ylab("Plastic Waste Per Capita")+
  xlab("Mismanaged Plastic Waste Per Captia")+
  geom_point()+
  facet_wrap(~continent)
```

### Exercise 7

There does not appear to be a relationship for either plot. 
```{r plastic-waste-population-total}
plastic_waste %>%filter(plastic_waste_per_cap < 3.5) %>% 
ggplot( mapping = aes(x = total_pop, 
                      y = plastic_waste_per_cap)) +
    labs(title ="Relationship between Total Population
and plastic waste per capita")+
  ylab("Plastic Waste Per Capita")+
  xlab("Total Population")+
  geom_point()
```

```{r plastic-waste-population-coastal}
plastic_waste %>%filter(plastic_waste_per_cap < 3.5) %>% 
ggplot( mapping = aes(x = coastal_pop, 
                      y = plastic_waste_per_cap)) +
    labs(title ="Relationship between Coastal Population and 
    Plastic Waste Per Capita")+
  ylab("Plastic Waste Per Capita")+
  xlab("Coastal Population")+
  geom_point()
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}

```

## Pro-Tips

### Excercise 3

Try this :D

ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
  
### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements



