---
title: "Lab 02 - Plastic waste"
author: "Conor Lacey"
date: "01/19/23"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

Histogram of plastic waste per capita faceted by continent

```{r plastic-waste-continent}
plastic_waste %>%filter(plastic_waste_per_cap < 3.5) %>% #filtering out 3.5
ggplot(aes(x = plastic_waste_per_cap)) +
  facet_wrap(~continent)+
  labs(title = "Plastic Waste Per Capita by Continent")+
  xlab("Plastic Waste Per Capita")+
  ylab("Count")+
  geom_histogram(binwidth = 0.1,color="black")
```

From the histogram it appears Africa is the least variant continent
of all the continents. It also appears that North America is the most variant 
continent of all the continents.

### Exercise 2

The density distribution of plastic waste per capita faceted by continent
```{r plastic-waste-density}
plastic_waste %>%filter(plastic_waste_per_cap < 3.5) %>% 
ggplot(aes(x = plastic_waste_per_cap,
           color=continent,
           fill=continent)) +
  labs(title ="Plastic waste per capita by continent")+
  xlab("Plastic Waste Per Capita")+
  ylab("Density")+
  geom_density(alpha=0.3)
```

### Exercise 3

We defined the color and fill of the curves by mapping the aesthetics because they are
based on the data values (the continent label). The alpha level is not based on the data values,therefore it goes in the geom as a characteristic. 

### Exercise 4

In contrast to box plots, violins reveal the continuous density. It is apparent here
that the violins are just a mirrored density distribution of each continent displayed 
in a similar way to box plots. 

The box plots are able to display the quartiles and the median of the data. Violins 
cannot do this.

```{r plastic-waste-violin}
plastic_waste %>%filter(plastic_waste_per_cap < 3.5) %>% 
ggplot( mapping = aes(x = continent, 
                      y = plastic_waste_per_cap)) +
    labs(title ="Plastic waste per capita by continent")+
  ylab("Plastic Waste Per Capita")+
  xlab("Continent")+
  geom_violin()
```

### Exercise 5

There appears to be a positive relationship here between plastic and mismanaged plastic 
waste per capita. However, it also appears this relationship weakens with increased levels of plastic waster per capita.
```{r plastic-waste-mismanaged}
plastic_waste %>%filter(plastic_waste_per_cap < 3.5) %>% 
ggplot( mapping = aes(x = plastic_waste_per_cap, 
                      y = mismanaged_plastic_waste_per_cap)) +
    labs(title ="Relationship between Plastic and Mismanaged Plastic Per Capita")+
  ylab("Mismanged Plastic Waste Per Capita")+
  xlab("Plastic Waste Per Captia")+
  geom_point()
```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}
# insert code here
```

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```

## Pro-Tips

### Excercise 3

Try this :D

ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
  
### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements



